### 2015-04-03 src/paper4/p4_discussion.tex
Updated paper 4 and figure in paper 3

In our previous experiment we compared translation perception with body-fixed versus world-fixed fixations at near depth (50 \si{\centi\metre}) only \cite{clemens2015a}. \figref{p4:fig6} shows how well our $\alpha_{50}$ parameter explains the data in our previous paper. The highly positive correlation between the actual PSEs in the previous study and those predicted using the model of the present paper ($\rho = 0.60$, $p = 0.06$) adds confidence to the parameter values presented here. The average difference between the values found here and those reported previously (see \tabref{p4:tab2}) is 12 \textpm 8 percent-points, which is relatively small given the indpendent measurements.

% Relation to the VOR
The function of the LVOR is to keep the eyes stable in the world during linear translation \cite{paige1989,busettini1994,paige1998}. Because it also needs to scale with fixation depth \cite{angelaki2004}, it is possible that both the LVOR and self-motion perception have the same underlying same signal. Because of the visual fixation point, visual following mechanisms may even augment the LVOR compensation. If the oculomotor signal generated by the LVOR is used for self-motion perception, one would expect that the LVOR compensation at 50 and 200cm closely relate to the corresponding oculomotor weights in the present study (see \figref{p4:fig5} and \tabref{p4:tab2}). To further explore this, we derived the LVOR gains for 50 and 200 \si{\centi\metre} from Paige et al. \citeyear{paige1989} and computed their expected depth ratio, $d_{200} / d_{50}$. This ratio, 1.87, is in between the ratio of the 6 participants who did not show any sign of scaling ($\frac{d_{200}}{d_{50}} = 1.07 \pm 0.16$) and the 2 participants that did show scaling ($\frac{d_{200}}{d_{50}} = 3.06 \pm 0.27$).

@@

It is important to point out that while the vestibular signal and thus noise is constant for a given translation distance, the noise in the associated oculomotor estimate might not be, as the magnitude of the eye movement (i.e. signal-dependent noise) is modulated by fixation depth. If signal-dependent noise would play a role, nearby world-fixed fixations would cause larger eye movements with more noise compared to the far world-fixed fixations with smaller eye movements. The oculomotor based translation estimate would be weighted less for near versus far fixation, potentially explaining the partial compensation for fixation depth we have observed. In addition, the noise levels in the oculomotor estimate could also depend on fixation depth itself: the retinal displacement of a world stationary fixation point decreases with fixation depth, making it less informative about the amount of self-motion. The noise level in the oculomotor estimate would therefore be higher for far away compared to nearby fixation points. For world stationary targets, this would predict an underestimation of self-motion while fixating far away compared to nearby, which is in line with our observations. However, it also predicts a similar effect for body stationary targets. As no such effects between the near and far body stationary fixation targets have been observed, we consider it an unlikely alternative explanation.

>>>>

In our previous experiment we compared translation perception with body-fixed versus world-fixed fixations at near depth (50 \si{\centi\metre}) only \cite{clemens2015a}. \figref{p4:fig6} shows how well our $\alpha_{50}$ parameter explains the data in our previous paper. The positive correlation between the actual PSEs in the previous study and those predicted using the model of the present paper ($\rho = 0.60$, $p = 0.06$) adds confidence to the parameter values presented here. The average difference between the values found here and those reported previously (see \tabref{p4:tab2}) is 12 \textpm 8 percent-points, which is relatively small given the indpendent measurements.

% Relation to the VOR
The function of the LVOR is to keep the eyes stable in the world during linear translation \cite{paige1989,busettini1994,paige1998}. Because it also needs to scale with fixation depth \cite{angelaki2004}, it is possible that the LVOR and self-motion perception have the same underlying signal. Because of the visual fixation point in our paradigm, visual following mechanisms may even augment the LVOR compensation. If the oculomotor signal generated by the LVOR is used for self-motion perception, one would expect that the LVOR compensation at 50 and 200cm closely relate to the corresponding oculomotor weights in the present study (see \figref{p4:fig5} and \tabref{p4:tab2}). To further explore this, we derived the LVOR gains for 50 and 200 \si{\centi\metre} from Paige et al. \citeyear{paige1989} and computed their expected depth ratio, $d_{200} / d_{50}$. This ratio, 1.87, is in between the ratio of the 6 participants who did not show any sign of scaling ($\frac{d_{200}}{d_{50}} = 1.07 \pm 0.16$) and the 2 participants that did show scaling ($\frac{d_{200}}{d_{50}} = 3.06 \pm 0.27$), suggesting that our effect might share a pathway with the LVOR.

@@

It is important to point out that while the vestibular signal and thus noise is constant for a given translation distance, the noise in the associated oculomotor estimate might change with fixation distance because the magnitude of the eye movement is modulated by fixation depth and may show signal-dependent noise.

If this were the case, the nearby world-fixed fixations would cause larger eye movements with more noise compared to the far world-fixed fixations with smaller eye movements. The oculomotor based translation estimate would be weighted less for near versus far fixation, potentially explaining the partial compensation for fixation depth we have observed. In addition, the noise levels in the oculomotor estimate could also depend on fixation depth itself: the retinal displacement of a world stationary fixation point decreases with fixation depth, making it less informative about the amount of self-motion. The noise level in the oculomotor estimate would therefore be higher for far away compared to nearby fixation points. For world stationary targets, this would predict an underestimation of self-motion while fixating far away compared to nearby, which is in line with our observations. However, it also predicts a similar effect for body stationary targets. As no such effects between the near and far body stationary fixation targets have been observed, we consider it an unlikely alternative explanation.

### 2015-04-03 src/paper4/p4_intro.tex
Updated paper 4 and figure in paper 3

An accurate internal estimate of self-motion is required to navigate effectively through a complex three-dimensional environment. The vestibular system as well as optic flow provide essential information about self-motion \cite{gibson1955, benson1986, harris2000, israel1989, angelaki2005, carriot2013, chen2010}. During navigation, however, the eye movements that typically occur to maintain visual acuity on important objects in the environment disturb the optic flow patterns. Using the oculomotor signal, the brain is able to account for these disturbances by internally separating optic flow in two components, one caused by self-motion and the other by eye movement \cite{warren1988, royden1992, freeman1998, lappe1999}.

@@

Geometrically, eye movements that keep fixation on a world-centered target during lateral whole body translation (i.e. the linear vestibulo-ocular reflex; LVOR), must scale with fixation depth \cite{angelaki2004}. When fixating body-centered targets these eye movements must be suppressed irrespective of fixation distance \cite{angelaki2004}. Conversely, to serve as an adequate self-motion cue, the brain must internally scale the ensuing eye movement by fixation distance. Because we did not manipulate fixation distance in our previous study, it is unknown whether eye movements are used as a rudimentary cue for self-motion (i.e. without taking fixation depth into account), or are properly scaled in the mechanisms for self-motion perception.

>>>>

An accurate internal estimate of self-motion is required to navigate effectively through a complex three-dimensional environment. The vestibular system as well as optic flow provide essential information about self-motion \cite{gibson1955, benson1986, harris2000, israel1989, angelaki2005, carriot2013, chen2010}. During navigation, however, the eyes typically move to maintain visual acuity on important objects. These eye movements disturb the optic flow patterns. Using the oculomotor signal, the brain is able to account for these disturbances by internally separating optic flow into two components, one caused by self-motion and the other by eye movement \cite{warren1988, royden1992, freeman1998, lappe1999}.

@@

Geometrically, eye movements that keep fixation on a world-centered target during lateral whole body translation (i.e. the linear vestibulo-ocular reflex; LVOR), must scale with fixation depth \cite{angelaki2004}. When fixating body-centered targets these eye movements must be suppressed irrespective of fixation distance \cite{angelaki2004}. Conversely, when fixating world-centered targets, the brain must internally scale the ensuing eye movement by fixation distance to serve as an adequate self-motion cue. Because we did not manipulate fixation distance in our previous study, we could not dissociate whether eye movements are used as a rudimentary cue for self-motion (i.e. without taking fixation depth into account), or are properly scaled in the mechanisms for self-motion perception.

### 2015-04-03 src/paper4/p4_results.tex
Updated paper 4 and figure in paper 3

In a recent study we have shown that eye movement signals contribute to the perception of body translation, even in the absence of optic flow or a visual fixation point \cite{clemens2015a}. Because eye rotations must be scaled by target depth ($\varphi d = T$) to serve as an adequate translation cue, we tested self-motion perception for near (50 \si{\centi\metre}) and far (200 \si{\centi\metre}) fixations. Participants were presented with two subsequent translations \figref{p4:fig1} while they kept fixation on a world- or body-stationary target that was presented either nearby or far away. After the two translation intervals, they had to judge whether the second translation was longer or shorter than the first.

@@

We normalized the eye movement data by taking the average ratio between the measured eye excursion and the geometrically required eye displacement were the target world stationary. \figref{p4:fig2}B shows that these normalized eye displacements are about zero during body fixed fixations (brown and orange) and close to one in the world fixed fixations (purple and pink data points), for all participants. The question is whether these eye movements are scaled by fixation depth in order to interpret them as linear self-motion cues.

@@

	\caption{\panelref{A} and \panelref{B} Psychometric curves (colored lines) and associated binned data (circles) for one participant. Circle size represents the amount of trials within the bin. Psychometric curves before collapsing across reference order are shown as gray lines. \panelref{A} Body-fixed condition (brown) while fixation was either near (dark) or far (light). \panelref{B} World-fixed condition (purple) while fixation was either near (dark) or far (light).
	\panelref{C} and \panelref{D} Individual and average points of subjective equality (PSEs). Color scheme matches panels A and B.
	}
	\label{p4:fig3}
\end{figure}

To answer this question, \figref{p4:fig3} illustrates psychophysical data on self-motion perception of a single participant for the two-fixation depths in both the body (\figref{p4:fig3}A) and world condition (\figref{p4:fig3}B). Lighter and darker colors indicate which fixation depth was the reference translation (see figure legend). The influence of fixation depth is characterized by a shift of the psychometric functions relative to the 10 \si{\centi\metre} reference translation (i.e. the PSE). For example, the rightward shift of the pink curve in \figref{p4:fig3}B, representing the world-condition, means that with a far target a longer translation (\about15 \si{\centi\metre}) was required to be perceived equivalent to a 10 \si{\centi\metre} reference translation with nearby fixation. Likewise, the leftward shift of the purple curve indicates that a shorter translation with near fixation (\about6 \si{\centi\metre}) is required to be perceived the same as the 10 \si{\centi\metre} reference translation with far fixation. Together, these opposite biases suggest that translations are perceived shorter for fixations further away. The brown and orange curves in \figref{p4:fig3}A do not show such shifts, indicating that fixation depth (i.e. near versus far) has no effect in absence of eye movements (i.e. when fixating a body-fixed target during the translation).

Similar results were obtained for all participants, as shown by the individual PSEs (right column of \figref{p4:fig3}). Statistical significance of the effects of fixation depth was evaluated by comparing PSEs for the two fixation depths using a paired t-test. PSEs  differed significantly between the two fixation depths in the world condition, $t(7) = 5.42$, $p < 0.01$ (\figref{p4:fig3}C), but not in the body condition, $t(7) = -1.17$, $p = 0.28$ (\figref{p4:fig3}D), confirming the single subject results (\figref{p4:fig3}A, B). Thus, increasing fixation depth does not influence self-motion perception during body-fixed fixations, but causes self-motion to be perceived as shorter during world-fixed fixations.

@@

We used a simple linear model to quantify to what extent eye movements are scaled by fixation depth in order to serve as a linear self-motion cue (see \nameref{p4:sec:methods}). This model describes the perceived translation as a weighted average of a vestibular estimate, equal to the actual translation, and an oculomotor-based estimate. The latter estimate depends on the eye excursion which is possibly scaled by fixation depth. We used two weighting parameters, $\alpha_{50}$ and $\alpha_{200}$, one per fixation depth (see \tabref{p4:tab2} for best-fit values). Using these parameters we predicted the PSEs, i.e. $m_p$ in \eqnref{p4:eq5}, and plotted them against the actually observed PSEs in \figref{p4:fig4}. The positive correlation ($\rho = 0.xx$, $p < 0.01$)\footnote{DONT FORGET TO COMPUTE THIS; NEED ACCESS TO MATLAB THOUGH} between observed and predicted PSEs shows that our simple model does reasonably well in predicting perceptual performance.

>>>>

In a recent study we have shown that eye movement signals contribute to the perception of body translation, even in the absence of optic flow or a visual fixation point \cite{clemens2015a}. Because eye rotations must be scaled by target depth ($\varphi d = T$) to serve as an adequate translation cue, we tested self-motion perception for near (50 \si{\centi\metre}) and far (200 \si{\centi\metre}) fixations. Participants were presented with two subsequent translations \figref{p4:fig1} while they kept fixation on a world- or body-stationary target that was presented either nearby or far away. After the two translation intervals, participants had to judge whether the second translation was longer or shorter than the first.

@@

We normalized the eye movement data by taking the average ratio between the measured eye excursion and the geometrically required eye displacement were the target world stationary. \figref{p4:fig2}B shows that these normalized eye displacements are about zero during body fixed fixations (brown and orange) and close to one in the world fixed fixations (purple and pink data points), for all participants. The question is whether these eye movements are inversely scaled by fixation depth in order to interpret them as linear self-motion cues.

@@

	\caption{\panelref{A} and \panelref{B} Psychometric curves (colored lines) and associated binned data (circles) for one participant. Circle size represents the amount of trials within the bin. Psychometric curves before collapsing across reference order are shown as gray lines. \panelref{A} World-fixed condition (purple) while fixation was either near (dark) or far (light). \panelref{B} Body-fixed condition (brown) while fixation was either near (dark) or far (light).
	\panelref{C} and \panelref{D} Individual and average points of subjective equality (PSEs). Color scheme matches panels A and B.
	}
	\label{p4:fig3}
\end{figure}

To investigate this question, \figref{p4:fig3} illustrates psychophysical data on self-motion perception of a single participant for the two-fixation depths in both the world (\figref{p4:fig3}A) and body condition (\figref{p4:fig3}B). Lighter and darker colors indicate which fixation depth was the reference translation (see figure legend). The influence of fixation depth is characterized by a shift of the psychometric functions relative to the 10 \si{\centi\metre} reference translation (i.e. the PSE). For example, the rightward shift of the pink curve in \figref{p4:fig3}A, representing the world-condition, means that with a far target a longer translation (\about15 \si{\centi\metre}) was required to be perceived equivalent to a 10 \si{\centi\metre} reference translation with nearby fixation. Likewise, the leftward shift of the purple curve indicates that a shorter translation with near fixation (\about6 \si{\centi\metre}) is required to be perceived the same as the 10 \si{\centi\metre} reference translation with far fixation. Together, these opposite biases suggest that translations are perceived shorter for fixations further away. For the body condition, no shift of the psychometric curves is visible (\figref{p4:fig3B}), indicating that fixation depth (i.e. near versus far) has no effect in absence of eye movements.

Similar results were obtained for all participants, as shown by the individual PSEs (right column of \figref{p4:fig3}). Statistical significance of the effects of fixation depth was evaluated by comparing PSEs for the two fixation depths using a paired t-test. PSEs  differed significantly between the two fixation depths in the world condition, $t(7) = 5.42$, $p < 0.01$ (\figref{p4:fig3}C), but not in the body condition, $t(7) = -1.17$, $p = 0.28$ (\figref{p4:fig3}D), confirming the single subject results (\figref{p4:fig3}A, B). Thus, increasing fixation depth does not influence self-motion perception during body-stationary fixations, but causes self-motion to be perceived as shorter during world-stationary fixations.

@@

We used a simple linear model to quantify to what extent eye movements are scaled by fixation depth in order to serve as a linear self-motion cue (see \nameref{p4:sec:methods}). This model describes the perceived translation distance as a weighted average of a vestibular estimate, equal to the actual translation, and an oculomotor-based estimate. The latter estimate depends on the eye excursion which should be scaled by fixation depth to serve as a valid cue. Our model contained two weighting parameters, $\alpha_{50}$ and $\alpha_{200}$, one per fixation depth (see \tabref{p4:tab2} for best-fit values). Using these parameters we predicted the PSEs, i.e. $m_p$ in \eqnref{p4:eq5}, and plotted them against the actually observed PSEs in \figref{p4:fig4}. The positive correlation ($\rho = 0.xxxxxxxxxxxxxxxxxxxx-need-to-setup-matlab-for-this-xxxxxxxxxxxxxxxxxxx$, $p < 0.01$) between observed and predicted PSEs shows that our simple model does reasonably well in predicting perceptual performance.
